---
title: "DS311 - R Lab Assignment"
author: "Zain"
date: "12/7/2022"
output:
  html_document:
    theme: united
    highlight: tango
    df_print: paged
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Assignment 1

* In this assignment, we are going to apply some of the build in data set in R for descriptive statistics analysis.
* To earn full grade in this assignment, students need to complete the coding tasks for each question to get the result.
* After finished all the questions, knit the document into HTML format for submission.

### Question 1

Using the **mtcars** data set in R, please answer the following questions.

```{r}
# Loading the data
data(mtcars)

# Head of the data set
head(mtcars)
```

a. Report the number of variables and observations in the data set.

```{r}
# Enter your code here!
dim(mtcars)
rows = dim(mtcars)[1]
columns = dim(mtcars)[2]
# Answer:
print(paste("There are total of",rows , "variables and",columns , "observations in this data set."))
```



b. Print the summary statistics of the data set and report how many discrete and continuous variables are in the data set.

```{r}
# Enter your code here!
summary(mtcars)

# Answer:
print("There are 11 discrete variables and 0 continuous variables in this data set.")
```



c. Calculate the mean, variance, and standard deviation for the variable **mpg** and assign them into variable names m, v, and s. Report the results in the print statement.

```{r}
# Enter your code here!
m = mean(mtcars$mpg)
s = sd(mtcars$mpg)
v = s^2
 print(paste("The average of Mile Per Gallon from this data set is ",round(m, 2)  , " with variance ",round(v, 2)  , " and standard deviation",round(s,2)  , "."))
```



d. Create two tables to summarize 1) average mpg for each cylinder class and 2) the standard deviation of mpg for each gear class.

```{r}
# Enter your code here!
mtcars$gear = as.character(mtcars$gear)

library(dplyr)
table.mean = mtcars %>% group_by(cyl) %>% summarise(mean = mean(mpg))
table.sd = mtcars%>%group_by(cyl) %>% summarise(standard_deviation=sd(mpg))
table.mean
table.sd

```



e. Create a crosstab that shows the number of observations belong to each cylinder and gear class combinations. The table should show how many observations given the car has 4 cylinders with 3 gears, 4 cylinders with 4 gears, etc. Report which combination is recorded in this data set and how many observations for this type of car.

```{r}
# Enter your code here!
xtabs(~cyl+gear,data=mtcars)
#table(mtcars$cyl,mtcars$gear)
print("The most common car type in this data set is car with 8 cylinders and 3 gears. There are total of 12 cars belong to this specification in the data set.")

```


*** 

### Question 2

Use different visualization tools to summarize the data sets in this question.

a. Using the **PlantGrowth** data set, visualize and compare the weight of the plant in the three separated group. Give labels to the title, x-axis, and y-axis on the graph. Write a paragraph to summarize your findings.

```{r}
# Load the data set
data("PlantGrowth")

# Head of the data set
head(PlantGrowth)

# Enter your code here!
library(ggplot2)
graph2 =  ggplot(PlantGrowth,aes(x=group,y=weight,color=group)) +geom_boxplot()
graph2

```

Result:

=> Report a paragraph to summarize your findings from the plot!



b. Using the **mtcars** data set, plot the histogram for the column **mpg** with 10 breaks. Give labels to the title, x-axis, and y-axis on the graph. Report the most observed mpg class from the data set.

```{r}
hist(mtcars$mpg,breaks=10,main="Histogram of the Mile Per Gallon for cars",xlab="Mile Per Gallon",ylab="Frequency",col="green")
print("Most of the cars in this data set are in the class of 15 mile per gallon.")

```



c. Using the **USArrests** data set, create a pairs plot to display the correlations between the variables in the data set. Plot the scatter plot with **Murder** and **Assault**. Give labels to the title, x-axis, and y-axis on the graph. Write a paragraph to summarize your results from both plots.

```{r}
# Load the data set
data("USArrests")
# Head of the data set
head(USArrests)
# Enter your code here!
plot(USArrests$Murder,USArrests$Assault,main = "Relationship between the murder and the assult from all the US arrest",xlab="Murder",ylab="Assault",col="green",pch=19)
```

Result:

From the graph, there is a postitive linear relationship between the Murder and the Assault in the US arrest.



***

### Question 3

Download the housing data set from www.jaredlander.com and find out what explains the housing prices in New York City.

Note: Check your working directory to make sure that you can download the data into the data folder. 

```{r, echo=FALSE}
# Load and clean the housing data set 
download.file(url='https://www.jaredlander.com/data/housing.csv',
              destfile='data/housing.csv', mode='wb')
housingData <- read.csv('data/housing.csv')
housingData <- subset(housingData, 
                      select = c("Neighborhood", "Market.Value.per.SqFt", "Boro", "Year.Built"))
housingData <- na.omit(housingData)                   
```

a. Create your own descriptive statistics and aggregation tables to summarize the data set and find any meaningful results between different variables in the data set.

```{r}
# Head of the cleaned data set
head(housingData)

# Enter your code here!
housingData.neighbor = housingData %>% group_by (Neighborhood) %>%
  summarise(meanmarketvalue=mean(Market.Value.per.SqFt),min=min(Market.Value.per.SqFt)
            ,max=max(Market.Value.per.SqFt),oldestyear=min(Year.Built)
            ,newestyear=max(Year.Built))
housingData.Boro =housingData %>% group_by (Boro) %>%
  summarise(meanmarketvalue=mean(Market.Value.per.SqFt),min=min(Market.Value.per.SqFt)
            ,max=max(Market.Value.per.SqFt),oldestyear=min(Year.Built)
            ,newestyear=max(Year.Built))
housingData.neighbor
housingData.Boro
```

b. Create multiple plots to demonstrates the correlations between different variables. Remember to label all axes and give title to each graph.

```{r}
# Enter your code here!
#ggplot(housingData,aes(y=Year.Built,x=Boro,col=Boro)) + geom_boxplot()
#ggplot(housingData,aes(y=Market.Value.per.SqFt,x=Boro,col=Boro)) + geom_boxplot()


#plot(housingData$Market.Value.per.SqFt,housingData$Year.Built,col=as.factor(housingData$Boro),pch=19)

#legend("bottomright",legend=unique(housingData$Boro),fill=as.factor(unique(housingData$Boro)))
#plot(housingData$Market.Value.per.SqFt,housingData$Year.Built,col=as.factor(housingData$Neighborhood),pch=19)
#cor(housingData$Market.Value.per.SqFt,housingData$Year.Built
```

c. Write a summary about your findings from this exercise.

From these graphs above, the market value per square feet in Manhattan is the largest among 5 places: Bronx, Brooklyn, Manhattan, Queens and Staten Island.
Also, the correlation between market value per square feet and the year built is -0.0955, this is really close to 0, which indicates that there seems to be really small relationship between the year house was built AND the market value per square feet.


